---
title: "R Notebook"
output: html_notebook
---

# 311 NYPD Data
```{r}

```

## 1. Introduction

### 1a. Motivation (Tim)
311 used by government to monitor efficiency and also for public to keep government agencies accountable. 
Cite Resources. 

Use Case:
- Operations for 311: 
  - allocate more resources for incidents for areas and time 
- Incident Patterns:
  - new homeowners, find which location is less noisy/more noisy fit their needs
  - police, identifying underlying crime patterns (crime hot-spots)
  - policy-makers, lots of complain about traffic, fix roads/traffic lights

### 1b. Main Questions (Everyone): 
1. **Operations/Efficiency:** 
- What areas/month/time of day/incident types are police most/least efficient? (WHY: allocate resources)
- What causes calls to be most/least efficient? (Connect with distance to police station channel data type, time of day/year, etc.)
- Instead of efficiency analyze when police actually took action vs. false alarm. 

2. **Incident Patterns:**
- At which locations do specific incident types occur the most? Most notable patterns? (incident type, false alarm/actual action)
- At what time of day do specific incident types occur the most? Most notable patterns?
- (Tim) How do incidents change throughout the months/year? e.g. What years do some incidents peak? (Tim will do an animation showing the incidents by month/year in D3 (month/yearly process), but maybe finding a most notable pattern)

### 1c. Team Contribution (Everyone)

(Write how each person contributed to the project)
1. Donghan He: 
2. Nico Winata: 
3. Charlene Luo:
4. Tim Kartawijaya:

## 2. Description of Data (Donghan)
- Source
- How it was collected in NYC Open Data
- Noteworthy Features

## 3. Analysis of Data Quality (Donghan)
- Data Quality 
- Graphs/Code 

## 4. Main Analysis

```{r}
library(tidyverse)
df = read_csv("cleaned_data.csv")
```

### 4a. Data Preprocessing (Donghan)
- Explain data cleaning process
- Mention Python Script cleaning

In analyzing the 311 Data, we will format these analyses in the form of questions we want answered. **Include approaches that didn't work and challenges** 

QUESTIONS:
1. **Operations/Efficiency:** 
- What areas/month/time of day/incident types are police most/least efficient? (WHY: allocate resources)
- What causes calls to be most/least efficient? (Connect with distance to police station channel data type, time of day/year, etc.)
- Instead of efficiency analyze when police actually took action vs. false alarm. 

2. **Incident Patterns:**
- At which locations do specific incident types occur the most? Most notable patterns? (incident type, false alarm/actual action)
- At what time of day do specific incident types occur the most? Most notable patterns?
- (Tim) How do incidents change throughout the months/year? e.g. What years do some incidents peak? (Tim will do an animation showing the incidents by month/year in D3 (month/yearly process), but maybe finding a most notable pattern)

### 4b. Police Efficiency (Nico/Charlene?)

Police efficiency also has a visible geographical pattern across New York City. As seen in the following two visualizations, note that the distance to closest station is at a minimum in the requests from Manhattan while request from parts of Staten Island and Brooklyn and Queens have much higher distances. A similar pattern appears in the processing time of the requests from these areas. Processing times in Manhattan tend to be lower than those from the Bronx, Staten Island, and parts of Brooklyn and Queens. This leads us to conclude that the distance to the closest station has a large effect on the processing time of police requests. This information could lead to increased city efforts to allocate resources to Staten Island and parts of Queens to help with the processing time in those areas. 

### 4c. Incident Patterns
By looking at the zip codes of various types of incident types, we were able to visualize patterns of various incident frequencies. Below, we observed that street condition complaints were least common in Manhattan and occurred most often in parts of southern queens and Brooklyn as well as eastern Bronx.  This could be due to better road conditions in Manhattan. 
 

Next, from the map below,  there is a clear region in Brooklyn and Queens that has a much higher frequency of traffic complaints than anywhere else in New York City. This information suggests that Brooklyn and Queens are areas that need the most work in terms of traffic regulation. The higher ratio of residential areas in these regions may be a factor in this. 

 

From the map below, we observed that noise complaints occur most often in the Bronx, Lower East Side, and Brooklyn. This would be interesting to overlay with demographic information about resident age and occupation. These regions may have the highest noise complaints because there are more young people making noise at night. This information would also be useful for people interested in moving into New York City. Knowing which regions receive the most noise complaints could determine the best quiet (or lively) place to live that suits them.


 


However, not all types of complaints show a clear geographical pattern for frequencies. Below, we have a map of complaints regarding urinating in public. As shown, there is no clear borough that has the most urination incidents. From this map, we can conclude that public urination does not change predictably along geographic regions. More information is needed in order to analyze the underlying causes of higher public urination frequencies. 

 
## 5. Executive Summary (Monday together)
Provide a short nontechnical summary of the most revealing findings of your analysis written for a nontechnical audience. The length should be approximately two pages (if we were using pages…) Take extra care to clean up your graphs, ensuring that best practices for presentation are followed

## 6. Interactive Component (Tim)

## 7. Conclusion (Monday together)