---
title: "R Notebook"
output: html_notebook
---

# 311 NYPD Data
```{r}

```

## 1. Introduction

### 1a. Motivation (Tim)
311 used by government to monitor efficiency and also for public to keep government agencies accountable. 
Cite Resources. 

Use Case:
- Operations for 311: 
  - allocate more resources for incidents for areas and time 
- Incident Patterns:
  - new homeowners, find which location is less noisy/more noisy fit their needs
  - police, identifying underlying crime patterns (crime hot-spots)
  - policy-makers, lots of complain about traffic, fix roads/traffic lights

### 1b. Main Questions (Everyone): 
1. **Operations/Efficiency:** 
- What areas/month/time of day/incident types are police most/least efficient? (WHY: allocate resources)
- What causes calls to be most/least efficient? (Connect with distance to police station channel data type, time of day/year, etc.)
- Instead of efficiency analyze when police actually took action vs. false alarm. 

2. **Incident Patterns:**
- At which locations do specific incident types occur the most? Most notable patterns? (incident type, false alarm/actual action)
- At what time of day do specific incident types occur the most? Most notable patterns?
- (Tim) How do incidents change throughout the months/year? e.g. What years do some incidents peak? (Tim will do an animation showing the incidents by month/year in D3 (month/yearly process), but maybe finding a most notable pattern)

### 1c. Team Contribution (Everyone)

(Write how each person contributed to the project)
1. Donghan He: 
2. Nico Winata: 
3. Charlene Luo:
4. Tim Kartawijaya:

## 2. Description of Data (Donghan)
- Source
- How it was collected in NYC Open Data
- Noteworthy Features

## 3. Analysis of Data Quality (Donghan)
- Data Quality 
- Graphs/Code 

## 4. Main Analysis

```{r}
library(tidyverse)
df = read_csv("cleaned_data.csv")
```

### 4a. Data Preprocessing (Donghan)
- Explain data cleaning process
- Mention Python Script cleaning

In analyzing the 311 Data, we will format these analyses in the form of questions we want answered. **Include approaches that didn't work and challenges** 

QUESTIONS:
1. **Operations/Efficiency:** 
- What areas/month/time of day/incident types are police most/least efficient? (WHY: allocate resources)
- What causes calls to be most/least efficient? (Connect with distance to police station channel data type, time of day/year, etc.)
- Instead of efficiency analyze when police actually took action vs. false alarm. 

2. **Incident Patterns:**
- At which locations do specific incident types occur the most? Most notable patterns? (incident type, false alarm/actual action)
- At what time of day do specific incident types occur the most? Most notable patterns?
- (Tim) How do incidents change throughout the months/year? e.g. What years do some incidents peak? (Tim will do an animation showing the incidents by month/year in D3 (month/yearly process), but maybe finding a most notable pattern)

### 4b. Police Efficiency (Nico/Charlene?)

Police efficiency also has a visible geographical pattern across New York City. As seen in the following two visualizations, note that the distance to closest station is at a minimum in the requests from Manhattan while request from parts of Staten Island and Brooklyn and Queens have much higher distances. A similar pattern appears in the processing time of the requests from these areas. Processing times in Manhattan tend to be lower than those from the Bronx, Staten Island, and parts of Brooklyn and Queens. This leads us to conclude that the distance to the closest station has a large effect on the processing time of police requests. This information could lead to increased city efforts to allocate resources to Staten Island and parts of Queens to help with the processing time in those areas. 

### 4c. Incident Patterns


## 5. Executive Summary (Monday together)
Provide a short nontechnical summary of the most revealing findings of your analysis written for a nontechnical audience. The length should be approximately two pages (if we were using pages…) Take extra care to clean up your graphs, ensuring that best practices for presentation are followed

## 6. Interactive Component (Tim)

## 7. Conclusion (Monday together)